---
title: "Phytoplankton Analysis"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 12, fig.height = 6)

library(tidyverse)
library(crosstalk)
library(plotly)
library(patchwork)
library(rstatix)
library(ggpubr)
library(confintr)

d = readxl::read_excel("output/data_both_projects_rolled_up.xlsx")

d = d |> 
  dplyr::mutate(month = factor(month, levels = month.abb))

col_scheme = ggplot2::scale_fill_manual(
  values = c("Cyanophyceae" = 'lightblue',
             "Chlorophyceae" = 'orange',
             "Bacillariophyceae" = 'red',
             "Dinophyceae" = 'darkblue',
             "Chryso- & Cryptophyceae" = 'lightyellow')
)

my_plot_theme = theme_minimal() + 
  theme(axis.text = element_text(size = 10),
        axis.title = ggtext::element_markdown(size = 13))

# Functions #
run_t_test_year_lab = function(dat,var_name){
  dat |> 
    purrr::map( ~ {
      number_rows = nrow(.x)
      tryCatch(
        expr = {
          t.test(.x[[paste0("DB.",var_name)]], .x[[paste0("LV.",var_name)]], paired = T) |> 
            broom::tidy() |> 
            dplyr::mutate(test_run = TRUE,
                          project = unique(.x$project), 
                          year = unique(.x$year),
                          DB_mean = round(mean(.x$DB.cells.ml),5),
                          LV_mean = round(mean(.x$LV.cells.ml),5),
                          number_rows = number_rows) |> 
            dplyr::select(test_run,project,year,DB_mean,LV_mean,dplyr::everything())},
        error = function(e) return(
          data.frame(
            test_run = FALSE,
            project = unique(.x$project), 
            year = unique(.x$year),
            number_rows = number_rows,
            DB_mean = round(mean(.x$DB.cells.l,na.rm=T),5),
            LV_mean = round(mean(.x$LV.cells.ml,na.rm=T),5)))
      )
    },.progress=T) |> 
    dplyr::bind_rows() |> 
    tidyr::as_tibble()
}

run_t_test_year_lab_class = function(dat,var_name){
  dat |> 
    purrr::map( ~ {
      number_rows = nrow(.x)
      tryCatch(
        expr = {
          t.test(.x[[paste0("DB.",var_name)]], .x[[paste0("LV.",var_name)]], paired = T) |> 
            broom::tidy() |> 
            dplyr::mutate(test_run = TRUE,
                          project = unique(.x$project), 
                          year = unique(.x$year),
                          class.name = unique(.x$class.name),
                          DB_mean = round(mean(.x$DB.cells.ml),5),
                          LV_mean = round(mean(.x$LV.cells.ml),5),
                          number_rows = number_rows) |> 
            dplyr::select(test_run,project,year,DB_mean,LV_mean,dplyr::everything())},
        error = function(e) return(
          data.frame(
            test_run = FALSE,
            project = unique(.x$project), 
            year = unique(.x$year),
            class.name = unique(.x$class.name),
            number_rows = number_rows,
            DB_mean = round(mean(.x$DB.cells.l,na.rm=T),5),
            LV_mean = round(mean(.x$LV.cells.ml,na.rm=T),5)))
      )
    },.progress=T) |> 
    dplyr::bind_rows() |> 
    tidyr::as_tibble()
}

run_t_test = function(dat,var_name){
  dat |> 
    purrr::map( ~ {
      number_rows = nrow(.x)
      tryCatch(
        expr = {
          t.test(.x[[paste0("DB.",var_name)]], .x[[paste0("LV.",var_name)]], paired = T) |> 
            broom::tidy() |> 
            dplyr::mutate(test_run = TRUE,
                          project = unique(.x$project), 
                          year = unique(.x$year),
                          month = unique(.x$month),
                          basin = unique(.x$basin),
                          DB_mean = round(mean(.x$DB.cells.ml),5),
                          LV_mean = round(mean(.x$LV.cells.ml),5),
                          number_rows = number_rows) |> 
            dplyr::select(test_run,project,year,month,basin,DB_mean,LV_mean,dplyr::everything())},
        error = function(e) return(
          data.frame(
            test_run = FALSE,
            project = unique(.x$project), 
            year = unique(.x$year),
            month = unique(.x$month),
            basin = unique(.x$basin),
            number_rows = number_rows,
            DB_mean = round(mean(.x$DB.cells.l,na.rm=T),5),
            LV_mean = round(mean(.x$LV.cells.ml,na.rm=T),5)))
      )
    },.progress=T) |> 
    dplyr::bind_rows() |> 
    tidyr::as_tibble()
}

run_t_test_class = function(dat,var_name){
  dat |> 
    purrr::map( ~ {
      number_rows = nrow(.x)
      tryCatch(
        expr = {
          t.test(.x[[paste0("DB.",var_name)]], .x[[paste0("LV.",var_name)]], paired = T) |> 
            broom::tidy() |> 
            dplyr::mutate(test_run = TRUE,
                          project = unique(.x$project), 
                          year = unique(.x$year),
                          month = unique(.x$month),
                          basin = unique(.x$basin),
                          class.name = unique(.x$class.name),
                          DB_mean = round(mean(.x$DB.cells.ml),5),
                          LV_mean = round(mean(.x$LV.cells.ml),5),
                          number_rows = number_rows) |> 
            dplyr::select(test_run,project,year,month,basin,class.name,DB_mean,LV_mean,dplyr::everything())},
        error = function(e) return(
          data.frame(
            test_run = FALSE,
            project = unique(.x$project), 
            year = unique(.x$year),
            month = unique(.x$month),
            basin = unique(.x$basin),
            class.name = unique(.x$class.name),
            number_rows = number_rows,
            DB_mean = round(mean(.x$DB.cells.l,na.rm=T),5),
            LV_mean = round(mean(.x$LV.cells.ml,na.rm=T),5)))
      )
    },.progress=T) |> 
    dplyr::bind_rows() |> 
    tidyr::as_tibble()
}
```

```{r create_plotting_functions}
make_plot_by_month = function(d,
                              vars,
                              var_label, 
                              plot_type = NA,
                              t_test_results = NULL){
  
  if(plot_type == 'classes_by_month'){
    d_summarised = d |> 
      tidyr::pivot_longer(cols = vars, names_to = "lab", values_to = 'y') |> 
      dplyr::mutate(lab = ifelse(stringr::str_detect(lab,"^DB"),"Darren","Lidija")) |> 
      dplyr::group_by(project, year, basin, month, class.name, lab) |> 
      dplyr::reframe(y = sum(y, na.rm=T))
    
    # t_test_for_plot = t_test_results |> 
    #               dplyr::filter(project == unique(d_summarised$project),
    #                             year == unique(d_summarised$year)) |> 
    #   dplyr::left_join(d_summarised |> 
    #                      dplyr::group_by(month,basin,lab) |> 
    #                      dplyr::reframe(max_y = max(y)))
    
    p = ggplot(d_summarised, 
               aes(x = month)) + 
      col_scheme +
      geom_col(aes(y = y, fill = class.name)) +
      # geom_text(aes(y = max(d_summarised$y), label = ifelse(p.value < 0.05, '*', '')),
      # data = t_test_for_plot) +
      labs(y = paste0("Phytoplankton ", var_label), fill = '',
           x = "Month") +
      facet_wrap(lab ~ basin, ncol = 2, strip.position = "right") +
      my_plot_theme +
      theme(legend.position = 'bottom') + 
      coord_cartesian(clip = 'off')
  } 
  if(plot_type == 'species_richness'){
    # Darren's data
    db_d = d |> 
      dplyr::select(project, year, basin, month, DB.cells.ml, DB.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(DB.cells.ml != 0, DB.mm3.l != 0, !is.na(DB.cells.ml), !is.na(DB.mm3.l)) |> 
      dplyr::count(project, year, basin, month) |> 
      dplyr::mutate(lab = 'Darren')
    
    # Lidija's data
    lv_d = d |> 
      dplyr::select(project, year, basin, month, LV.cells.ml, LV.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(LV.cells.ml != 0, LV.mm3.l != 0, !is.na(LV.cells.ml), !is.na(LV.mm3.l)) |> 
      dplyr::count(project, year, basin, month) |> 
      dplyr::mutate(lab = 'Lidija')
    
    d_for_plot = dplyr::bind_rows(db_d, lv_d)#|> 
    # dplyr::filter(project == proj)
    
    p = ggplot(d_for_plot, 
               aes(x = month, y = n)) + 
      geom_point() + 
      scale_y_continuous(limits = c(0, max(d_for_plot$n)), breaks = seq(0,50,5)) +
      labs(y = paste0("Phytoplankton ", var_label),
           x = "Month") +
      facet_wrap(lab ~ basin, ncol = 2, strip.position = "right") +
      theme(legend.position = 'bottom') + 
      my_plot_theme
  }
  if(plot_type == 'edibility'){
    
    # Darren's data
    db_d = d |> 
      dplyr::select(project, year, basin, month, edibility, DB.cells.ml, DB.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(DB.cells.ml != 0, DB.mm3.l != 0, !is.na(DB.cells.ml), !is.na(DB.mm3.l)) |> 
      dplyr::count(project, year, basin, month, edibility) |> 
      dplyr::mutate(lab = 'Darren')
    
    # Lidija's data
    lv_d = d |> 
      dplyr::select(project, year, basin, month, 
                    edibility, LV.cells.ml, LV.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(LV.cells.ml != 0, LV.mm3.l != 0, !is.na(LV.cells.ml), !is.na(LV.mm3.l)) |> 
      dplyr::count(project, year, basin, month, edibility) |> 
      dplyr::mutate(lab = 'Lidija')
    
    d_for_plot = dplyr::bind_rows(db_d, lv_d)
    
    p = ggplot(d_for_plot, 
               aes(x = month, y = n, fill = lab)) + 
      geom_col(position = position_dodge()) + 
      scale_y_continuous(limits = c(0, max(d_for_plot$n)), breaks = seq(0,50,5)) +
      labs(y = paste0("Phytoplankton Species Counts by ", var_label),
           x = "Month") +
      facet_wrap(edibility ~ basin, ncol = 2, strip.position = "right") +
      my_plot_theme + 
      theme(legend.position = 'bottom') + 
      scale_fill_brewer(palette = "Set2")
  }
  if(plot_type == 'edibility_perc_diff'){
    
    # Darren's data
    db_d = d |> 
      dplyr::select(project, year, basin, month, edibility, DB.cells.ml, DB.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(DB.cells.ml != 0, DB.mm3.l != 0, !is.na(DB.cells.ml), !is.na(DB.mm3.l)) |> 
      dplyr::count(project, year, basin, month, edibility) |> 
      dplyr::mutate(lab = 'Darren')
    
    # Lidija's data
    lv_d = d |> 
      dplyr::select(project, year, basin, month, 
                    edibility, LV.cells.ml, LV.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(LV.cells.ml != 0, LV.mm3.l != 0, !is.na(LV.cells.ml), !is.na(LV.mm3.l)) |> 
      dplyr::count(project, year, basin, month, edibility) |> 
      dplyr::mutate(lab = 'Lidija')
    
    d_for_plot = dplyr::bind_rows(db_d, lv_d) |>
      tidyr::pivot_wider(names_from = lab, values_from = n, values_fill = 0) |> 
      dplyr::mutate(perc_diff = 100*(Darren - Lidija)/((Darren + Lidija)/2)) |> 
      dplyr::mutate(perc_diff = ifelse(Darren == 0 | Lidija == 0, NA, perc_diff)) |> 
      dplyr::mutate(`DB compared to LV` = ifelse(perc_diff > 0, "DB larger","LV larger")) |> 
      dplyr::filter(!is.na(perc_diff))
    
    p = ggplot(d_for_plot, 
               aes(x = month, y = perc_diff)) + 
      geom_col(position = position_dodge(), aes(fill = `DB compared to LV`)) + 
      labs(y = paste0("Percent Difference in Phytoplankton Species Counts by ", var_label),
           x = "Month") +
      facet_wrap(edibility ~ basin, ncol = 2, strip.position = "right") +
      my_plot_theme + 
      theme(legend.position = 'bottom') + 
      scale_fill_brewer(palette = "Set2")
  }
  p
}

```

# Sections {.tabset}

## Stats {.tabset}

### Descriptive Statistics

```{r}
# d |> 
#   tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l)) |> 
#   dplyr::group_by(name) |> 
#   dplyr::reframe(min_value = min(value),
#                  max_value = max(value),
#                  median_value = median(value),
#                  mean_value = mean(value)) |> 
#   dplyr::arrange(dplyr::desc(max_value)) |> 
#   dplyr::rename(variable = name) |> 
#   knitr::kable()
```

The following table shows means, plus or minus 95% confidence intervals, for abundance and biovolume for each lab separately.

```{r descriptive_stats_by_class_and_lab}
d |> 
  tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l)) |> 
  dplyr::mutate(value = round(value, 4)) |> 
  dplyr::group_by(name, class.name) |> 
  dplyr::reframe(mean_value = round(mean(value,na.rm=T),3),
                 low_cf_int = round(ci_mean(value)[[2]][1],3),
                 upper_cf_int = round(ci_mean(value)[[2]][2],3)) |> 
  dplyr::mutate(mean_w_CI = paste0(mean_value, " (95%CI: ",low_cf_int," - ",upper_cf_int,")")) |> 
  dplyr::select(name,class.name,mean_w_CI) |> 
  dplyr::mutate(lab = ifelse(stringr::str_detect(name,"^DB"),"DB","LV")) |>
  dplyr::mutate(variable = ifelse(stringr::str_detect(name,"cells"),"abundance","biovolume")) |>
  dplyr::mutate(variable = paste0(variable, "_",lab)) |> 
  dplyr::select(-c(name,lab)) |>  
  tidyr::pivot_wider(names_from = variable, values_from = mean_w_CI) |> 
  dplyr::select(class.name,abundance_DB,abundance_LV,dplyr::everything()) |> 
  knitr::kable()
```

We see some fairly large differences in the means and their confidence intervals, but to test if they are significantly difference, I think we need to split the data by project, year, and basin, in addition to class name. See the 'Pairwise T-test' tab for these results.

On average, what 'correction factor' (mean ratio) could we apply to Lidija's results such that they would mirror Darren's?

```{r}
d_for_mr = d |> 
  dplyr::filter(DB.cells.ml > 0 & LV.cells.ml > 0) |> 
  dplyr::select(DB.cells.ml, LV.cells.ml) |> 
  dplyr::mutate(db_div_by_lv = DB.cells.ml / LV.cells.ml) |> 
  dplyr::filter(db_div_by_lv < 100)

labA = d_for_mr$DB.cells.ml
labB = d_for_mr$LV.cells.ml

mean_ratio <- round(mean(labA / labB),2)
mean_ratio

```

-   compare each row of data for which Darren and Lidija both have reported cell counts for a given class.
-   only include rows where Darren's counts aren't more than 100 times as large as Lidija's
-   divide Darren's count by Lidija's count for each row, then take the average of those ratios (i.e. calculate the *mean ratio*), we get `r mean_ratio`. This means that we would need to multiple Lidija's counts by this correction factor to approximate Darren's counts.

To get more granular with this, let's plot Lidijia's data on Darren's data in a linear model.
```{r}
model_origin <- round(coef(lm(labB ~ 0 + labA)),2)
```

By using this linear model (with an intercept of 0, i.e. if Darren reports 0 then logically Lidija *should* be reporting 0), we find the slope of the linear model to be `r model_origin`; this linear model suggests that we should multiple Lidija's data by roughly `r model_origin` to approximate Darren's results.

The reason we're getting such large "correction factors" seems to be that Darren sometimes has counts that are much larger than Lidija's. See histogram below (for this histogram, I have NOT excluded rows where Darren's counts are more than 100 times as large as Lidija's - I did that for the calculations above)

```{r}
d |> 
  dplyr::filter(DB.cells.ml > 0 & LV.cells.ml > 0) |> 
  dplyr::select(DB.cells.ml, LV.cells.ml) |> 
  dplyr::mutate(DB_div_by_LV = DB.cells.ml / LV.cells.ml) |> 
  ggplot() + 
  geom_histogram(aes(DB_div_by_LV))
```

### Abundance over Time by Class

#### Wahleach

```{r fig.height=12}
2023:2024 |> 
  purrr::map( ~ {
    c("NB","SB") |> 
      lapply(\(y){
        d |> 
          dplyr::filter(project == 'Wahleach') |>
          dplyr::filter(year == .x) |> 
          dplyr::filter(basin == y) |> 
          tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l), names_to = "lab") |> 
          dplyr::filter(lab %in% c("DB.cells.ml","LV.cells.ml")) |> 
          dplyr::group_by(project, year, month, basin, class.name, lab) |> 
          dplyr::reframe(value = sum(value,na.rm=T)) |> 
          dplyr::ungroup() |> 
          ggplot(aes(x = month, y = value, col = class.name, group = class.name)) + 
          geom_point() + 
          geom_line() +
          labs(title = paste0("Wahleach, ",.x,", ",y), y = "Abundance") +
          facet_wrap(class.name ~ lab, scales = 'free', ncol = 2) + 
          my_plot_theme +
          theme(legend.position = 'bottom')
      })
  })
```

#### Alouette

```{r fig.height=12}
2023:2024 |> 
  purrr::map( ~ {
    c("NB","SB") |> 
      lapply(\(y){
        d |> 
          dplyr::filter(project == 'Alouette') |>
          dplyr::filter(year == .x) |> 
          dplyr::filter(basin == y) |> 
          tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l), names_to = "lab") |> 
          dplyr::filter(lab %in% c("DB.cells.ml","LV.cells.ml")) |> 
          dplyr::group_by(project, year, month, basin, class.name, lab) |> 
          dplyr::reframe(value = sum(value,na.rm=T)) |> 
          dplyr::ungroup() |> 
          ggplot(aes(x = month, y = value, col = class.name, group = class.name)) + 
          geom_point() + 
          geom_line() +
          labs(title = paste0("Alouette, ",.x,", ",y), y = "Abundance") +
          facet_wrap(class.name ~ lab, scales = 'free', ncol = 2) + 
          my_plot_theme +
          theme(legend.position = 'bottom')
      })
  })
```

### Edibility over Time

#### Wahleach

```{r fig.height=12}
2023:2024 |> 
  purrr::map( ~ {
    c("NB","SB") |> 
      lapply(\(y){
        d |> 
          dplyr::filter(project == 'Wahleach') |>
          dplyr::filter(year == .x) |> 
          dplyr::filter(basin == y) |> 
          tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l), names_to = "lab") |> 
          dplyr::filter(lab %in% c("DB.cells.ml","LV.cells.ml")) |> 
          dplyr::group_by(project, year, month, basin, edibility, lab) |> 
          dplyr::reframe(value = sum(value,na.rm=T)) |> 
          dplyr::ungroup() |> 
          ggplot(aes(x = month, y = value, col = edibility, group = edibility)) + 
          geom_point() + 
          geom_line() +
          labs(title = paste0("Wahleach, ",.x,", ",y), y = "Abundance") +
          facet_wrap(edibility ~ lab, scales = 'free', ncol = 2) + 
          my_plot_theme +
          theme(legend.position = 'bottom')
      })
  })
```

#### Alouette

```{r fig.height=12}
2023:2024 |> 
  purrr::map( ~ {
    c("NB","SB") |> 
      lapply(\(y){
        d |> 
          dplyr::filter(project == 'Alouette') |>
          dplyr::filter(year == .x) |> 
          dplyr::filter(basin == y) |> 
          tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l), names_to = "lab") |> 
          dplyr::filter(lab %in% c("DB.cells.ml","LV.cells.ml")) |> 
          dplyr::group_by(project, year, month, basin, edibility, lab) |> 
          dplyr::reframe(value = sum(value,na.rm=T)) |> 
          dplyr::ungroup() |> 
          ggplot(aes(x = month, y = value, col = edibility, group = edibility)) + 
          geom_point() + 
          geom_line() +
          labs(title = paste0("Alouette, ",.x,", ",y), y = "Abundance") +
          facet_wrap(edibility ~ lab, scales = 'free', ncol = 2) + 
          my_plot_theme +
          theme(legend.position = 'bottom')
      })
  })
```

### Pairwise T-test {.tabset}

```{r perform_paired_t_test}
#1.	Can you look to see if there are significant differences between the larger species between the 2 labs each month and each year (for example, cyanophyceae, Chlorophyceae, etc.) and I’m wondering if you could put a table together of the % differences?

# This pairing of t-tests is for Abundance
abund_lab_paired_t_test_results_within_class =  d |> 
  dplyr::group_by(project, year, month, basin, class.name) |> 
  dplyr::group_split() |> 
  run_t_test_class('cells.ml')

# Also do this where we put all classes together inside each month.
abund_lab_paired_t_test_results = d |> 
  dplyr::group_by(project, year, month, basin) |> 
  dplyr::group_split() |> 
  run_t_test('cells.ml')

# This pairing of t-tests is for Biovolume.
biov_lab_paired_t_test_results_within_class = d |> 
  dplyr::group_by(project, year, month, basin, class.name) |> 
  dplyr::group_split() |> 
  run_t_test_class('mm3.l')

# Also do this where we put all classes together inside each month.
biov_lab_paired_t_test_results = d |> 
  dplyr::group_by(project, year, month, basin) |> 
  dplyr::group_split() |> 
  run_t_test('mm3.l')

# Also look at abundance by year and lab
abund_proj_year_lab_paired_t_test_results = d |> 
  dplyr::group_by(project, year) |> 
  dplyr::group_split() |> 
  run_t_test_year_lab('cells.ml')

# Also look at abundance by year and lab
abund_proj_year_lab_class_paired_t_test_results = d |> 
  dplyr::group_by(project, year, class.name) |> 
  dplyr::group_split() |> 
  run_t_test_year_lab_class('cells.ml')

abund_lab_paired_t_test_box_plots = abund_lab_paired_t_test_results_within_class |> 
  dplyr::filter(p.value < 0.05) |> 
  dplyr::mutate(sig_p_value = p.value <= 0.05) |> 
  dplyr::group_by(project, year, month, basin, sig_p_value) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    d |> 
      dplyr::filter(year == .x$year, month == .x$month,
                    basin == .x$basin, project == .x$project) |> 
      dplyr::select(-c(DB.mm3.l,LV.mm3.l)) |> 
      tidyr::pivot_longer(cols = c(DB.cells.ml,LV.cells.ml)) |>
      dplyr::filter(class.name %in% na.omit(.x[.x$sig_p_value,]$class.name)) |> 
      ggplot(aes(x = class.name, y = value, fill = name)) + 
      geom_boxplot() + 
      labs(x = "Class", y = "Cell Count", fill = "Lab",
           title = paste0("Classes with Significantly Different Cell Numbers between Labs"),
           subtitle = paste0(.x$month," ",.x$year,", ",.x$basin,", ",.x$project)) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  })


biov_lab_paired_t_test_box_plots = biov_lab_paired_t_test_results_within_class |> 
  dplyr::filter(p.value < 0.05) |> 
  dplyr::mutate(sig_p_value = p.value <= 0.05) |> 
  dplyr::group_by(project, year, month, basin, class.name, sig_p_value) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    d |> 
      dplyr::filter(year == .x$year, month == .x$month,
                    basin == .x$basin, project == .x$project) |> 
      dplyr::select(-c(DB.cells.ml,LV.cells.ml)) |> 
      tidyr::pivot_longer(cols = c(DB.mm3.l,LV.mm3.l)) |> 
      dplyr::filter(class.name %in% na.omit(.x[.x$sig_p_value,]$class.name)) |> 
      ggplot(aes(x = class.name, y = value, fill = name)) + 
      geom_boxplot() + 
      labs(x = "Class", y = "Biovolume", fill = "Lab",
           title = paste0("Classes with Significantly Different Biovolume between Labs"),
           subtitle = paste0(.x$month," ",.x$year,", ",.x$basin,", ",.x$project)) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  })
```

#### Table {.tabset}

Note 1: These T-tests compare the Abundance andBiovolume recordings of Darren and Lidija by project, year, month, basin; an additional pair of tests also split the data by class name. Some of these data 'cuts' had too few rows of data to run the test (e.g. a single row): see column "number.rows" to see the total number of rows of data for both labs combined.

Note 2: percent difference of mean Biovolume is calculated as $100*(A-B)/((A+B)/2)$, where *A* is Darren's lab's values (`DB.mm3.l`) and *B* is Lidija's lab's values (`LV.mm3.l`); this means that Lidija's values are the 'reference' values, and Darren's values are displayed relative to those.

##### T-test by year and lab

```{r}
abund_proj_year_lab_paired_t_test_results |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

##### T-test by year, lab and class name

```{r}
abund_proj_year_lab_class_paired_t_test_results |> 
  dplyr::filter(!is.na(class.name)) |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

##### T-Test for Classes Separately

[Abundance:]{.underline}

```{r}
abund_lab_paired_t_test_results_within_class |> 
  dplyr::filter(!is.na(class.name)) |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

[Biovolume:]{.underline}

```{r}
biov_lab_paired_t_test_results_within_class |> 
  dplyr::filter(!is.na(class.name)) |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

##### T-Tests by Month (Classes Combined within each month)

[Abundance:]{.underline}

```{r}
abund_lab_paired_t_test_results |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

[Biovolume:]{.underline}

```{r}
biov_lab_paired_t_test_results |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

#### Boxplots

Just taking the statistically significantly different means, let's visualize them as boxplots. See the plot subtitles for details of exactly which cuts of the data are being visualized.

```{r}
abund_lab_paired_t_test_box_plots
```

```{r}
biov_lab_paired_t_test_box_plots
```

## Lab Comparison Plots {.tabset}

### Wahleach {.tabset}

#### Abundance by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Abundance by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Biovolume by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Biovolume by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Species Richness by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Species Richness by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Edibility by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Edibility by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Percent Difference Edibility by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility_perc_diff',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Percent Difference Edibility by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility_perc_diff',
                     t_test_results = abund_lab_paired_t_test_results)
```

### Alouette {.tabset}

#### Abundance by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Abundance by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Biovolume by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Biovolume by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Species Richness by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Species Richness by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Edibility by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Edibility by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Percent Difference Edibility by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility_perc_diff',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Percent Difference Edibility by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility_perc_diff',
                     t_test_results = abund_lab_paired_t_test_results)
```
