---
title: "Phytoplankton Analysis"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, fig.width = 12, fig.height = 6)

library(tidyverse)
library(crosstalk)
library(plotly)
library(patchwork)
library(rstatix)
library(ggpubr)

d = readxl::read_excel("output/data_both_projects_rolled_up.xlsx")

d = d |> 
  dplyr::mutate(month = factor(month, levels = month.abb))

col_scheme = ggplot2::scale_fill_manual(
  values = c("Cyanophyceae" = 'lightblue',
             "Chlorophyceae" = 'orange',
             "Bacillariophyceae" = 'red',
             "Dinophyceae" = 'darkblue',
             "Chryso- & Cryptophyceae" = 'lightyellow')
)

my_plot_theme = theme_minimal() + 
  theme(axis.text = element_text(size = 10),
        axis.title = ggtext::element_markdown(size = 13))
```

```{r create_plotting_functions}
make_plot_by_month = function(d,
                              vars,
                              var_label, 
                              plot_type = NA,
                              t_test_results = NULL){
  
  if(plot_type == 'classes_by_month'){
    d_summarised = d |> 
      tidyr::pivot_longer(cols = vars, names_to = "lab", values_to = 'y') |> 
      dplyr::mutate(lab = ifelse(stringr::str_detect(lab,"^DB"),"Darren","Lidija")) |> 
      dplyr::group_by(project, year, basin, month, class.name, lab) |> 
      dplyr::reframe(y = sum(y, na.rm=T))
    
    # t_test_for_plot = t_test_results |> 
    #               dplyr::filter(project == unique(d_summarised$project),
    #                             year == unique(d_summarised$year)) |> 
    #   dplyr::left_join(d_summarised |> 
    #                      dplyr::group_by(month,basin,lab) |> 
    #                      dplyr::reframe(max_y = max(y)))
    
    p = ggplot(d_summarised, 
               aes(x = month)) + 
      col_scheme +
      geom_col(aes(y = y, fill = class.name)) +
      # geom_text(aes(y = max(d_summarised$y), label = ifelse(p.value < 0.05, '*', '')),
      # data = t_test_for_plot) +
      labs(y = paste0("Phytoplankton ", var_label), fill = '',
           x = "Month") +
      facet_wrap(lab ~ basin, ncol = 2, strip.position = "right") +
      my_plot_theme +
      theme(legend.position = 'bottom') + 
      coord_cartesian(clip = 'off')
  } 
  if(plot_type == 'species_richness'){
    # Darren's data
    db_d = d |> 
      dplyr::select(project, year, basin, month, DB.cells.ml, DB.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(DB.cells.ml != 0, DB.mm3.l != 0, !is.na(DB.cells.ml), !is.na(DB.mm3.l)) |> 
      dplyr::count(project, year, basin, month) |> 
      dplyr::mutate(lab = 'Darren')
    
    # Lidija's data
    lv_d = d |> 
      dplyr::select(project, year, basin, month, LV.cells.ml, LV.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(LV.cells.ml != 0, LV.mm3.l != 0, !is.na(LV.cells.ml), !is.na(LV.mm3.l)) |> 
      dplyr::count(project, year, basin, month) |> 
      dplyr::mutate(lab = 'Lidija')
    
    d_for_plot = dplyr::bind_rows(db_d, lv_d)#|> 
    # dplyr::filter(project == proj)
    
    p = ggplot(d_for_plot, 
               aes(x = month, y = n)) + 
      geom_point() + 
      scale_y_continuous(limits = c(0, max(d_for_plot$n)), breaks = seq(0,50,5)) +
      labs(y = paste0("Phytoplankton ", var_label),
           x = "Month") +
      facet_wrap(lab ~ basin, ncol = 2, strip.position = "right") +
      theme(legend.position = 'bottom') + 
      my_plot_theme
  }
  if(plot_type == 'edibility'){
    
    # Darren's data
    db_d = d |> 
      dplyr::select(project, year, basin, month, edibility, DB.cells.ml, DB.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(DB.cells.ml != 0, DB.mm3.l != 0, !is.na(DB.cells.ml), !is.na(DB.mm3.l)) |> 
      dplyr::count(project, year, basin, month, edibility) |> 
      dplyr::mutate(lab = 'Darren')
    
    # Lidija's data
    lv_d = d |> 
      dplyr::select(project, year, basin, month, 
                    edibility, LV.cells.ml, LV.mm3.l, spp) |> 
      # Ensure rows we're keeping rows where DB reported values
      dplyr::filter(LV.cells.ml != 0, LV.mm3.l != 0, !is.na(LV.cells.ml), !is.na(LV.mm3.l)) |> 
      dplyr::count(project, year, basin, month, edibility) |> 
      dplyr::mutate(lab = 'Lidija')
    
    d_for_plot = dplyr::bind_rows(db_d, lv_d)
    
    # t_test_for_plot = t_test_results |> 
    #               dplyr::filter(project == unique(d_for_plot$project),
    #                             year == unique(d_for_plot$year)) |> 
    #   dplyr::left_join(d_for_plot |> 
    #                      dplyr::group_by(month,basin,lab) |> 
    #                      dplyr::reframe(max_y = max(n)))
    
    # browser()
    p = ggplot(d_for_plot, 
               aes(x = month, y = n, fill = lab)) + 
      geom_col(position = position_dodge()) + 
      # geom_text(aes(y = max(d_for_plot$n), label = ifelse(p.value < 0.05, '*', '')),
      # data = t_test_for_plot) +
      scale_y_continuous(limits = c(0, max(d_for_plot$n)), breaks = seq(0,50,5)) +
      labs(y = paste0("Phytoplankton Species Counts by ", var_label),
           x = "Month") +
      facet_wrap(edibility ~ basin, ncol = 2, strip.position = "right") +
      my_plot_theme + 
      theme(legend.position = 'bottom') + 
      scale_fill_brewer(palette = "Set2")
  }
  p
}

```

# Sections {.tabset}

## Stats {.tabset}

### Descriptive Statistics

```{r}
d |> 
  tidyr::pivot_longer(cols = c(LV.cells.ml,LV.mm3.l,DB.cells.ml,DB.mm3.l)) |> 
  dplyr::mutate(value = round(value, 4)) |> 
  dplyr::group_by(name) |> 
  dplyr::reframe(min_value = min(value),
                 max_value = max(value),
                 median_value = median(value),
                 mean_value = mean(value)) |> 
  dplyr::arrange(dplyr::desc(max_value)) |> 
  dplyr::rename(variable = name) |> 
  knitr::kable()
```

### Pairwise T-test {.tabset}

```{r perform_paired_t_test}
#1.	Can you look to see if there are significant differences between the larger species between the 2 labs each month and each year (for example, cyanophyceae, Chlorophyceae, etc.) and I’m wondering if you could put a table together of the % differences?

# biov_by_class = d |> 
#   dplyr::filter(year == 2023, project == 'Alouette') |> 
#   dplyr::select(-c(edibility,DB.cells.ml,LV.cells.ml)) |> 
#   tidyr::pivot_longer(cols = c(DB.mm3.l, LV.mm3.l)) |> 
#   dplyr::mutate(lab = ifelse(stringr::str_detect(name,"^DB"),"Darren","Lidija")) |> 
#   dplyr::group_by(project, year, month, basin, lab, class.name) |>
#   # Find the sum of biovolume per class name; keep the top two.
#   dplyr::mutate(biov_of_class = sum(value,na.rm=T)) |> 
#   dplyr::arrange(project,year,month,basin,lab,class.name,dplyr::desc(biov_of_class)) |> 
#   dplyr::ungroup()

# paired t-test by class.name and lab.
# Trying it out for 2023, Apr, Flagellates
# d_t = d |> 
#   dplyr::filter(project == 'Alouette', year == 2023, month == 'Apr', basin == 'NB') |> 
#   dplyr::mutate(the_row = dplyr::row_number()) |> 
#   dplyr::select(-c(DB.cells.ml,LV.cells.ml))

# This pairing of t-tests is for Abundance
abund_lab_paired_t_test_results_within_class = d |> 
  dplyr::group_by(project, year, month, basin, class.name) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    tryCatch(
      expr = {
        t.test(.x$DB.cells.ml, .x$LV.cells.ml, paired = T) |> 
          broom::tidy() |> 
          dplyr::mutate(test_run = TRUE,
                        project = unique(.x$project), 
                        year = unique(.x$year),
                        month = unique(.x$month),
                        basin = unique(.x$basin),
                        class.name = unique(.x$class.name),
                        DB_mean = round(mean(.x$DB.cells.ml),5),
                        LV_mean = round(mean(.x$LV.cells.ml),5)) |> 
          dplyr::rename(number.rows = parameter) |> 
          dplyr::select(test_run,project,year,month,basin,class.name,DB_mean,LV_mean,dplyr::everything())},
      error = function(e) return(
        data.frame(
          test_run = FALSE,
          project = unique(.x$project), 
          year = unique(.x$year),
          month = unique(.x$month),
          basin = unique(.x$basin),
          class.name = unique(.x$class.name),
          number.rows = nrow(.x),
          DB_mean = round(mean(.x$DB.cells.l,na.rm=T),5),
          LV_mean = round(mean(.x$LV.cells.ml,na.rm=T),5)))
    )
  },.progress=T) |> 
  dplyr::bind_rows() |> 
  tidyr::as_tibble()

# Also do this where we put all classes together inside each month.
abund_lab_paired_t_test_results = d |> 
  dplyr::group_by(project, year, month, basin) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    tryCatch(
      expr = {
        t.test(.x$DB.cells.ml, .x$LV.cells.ml, paired = T) |> 
          broom::tidy() |> 
          dplyr::mutate(test_run = TRUE,
                        project = unique(.x$project), 
                        year = unique(.x$year),
                        month = unique(.x$month),
                        basin = unique(.x$basin),
                        DB_mean = round(mean(.x$DB.cells.ml),5),
                        LV_mean = round(mean(.x$LV.cells.ml),5)) |> 
          dplyr::rename(number.rows = parameter) |> 
          dplyr::select(test_run,project,year,month,basin,DB_mean,LV_mean,dplyr::everything())},
      error = function(e) return(
        data.frame(
          test_run = FALSE,
          project = unique(.x$project), 
          year = unique(.x$year),
          month = unique(.x$month),
          basin = unique(.x$basin),
          number.rows = nrow(.x),
          DB_mean = round(mean(.x$DB.cells.ml,na.rm=T),5),
          LV_mean = round(mean(.x$LV.cells.ml,na.rm=T),5)))
    )
  },.progress=T) |> 
  dplyr::bind_rows() |> 
  tidyr::as_tibble()

# This pairing of t-tests is for Biovolume.
biov_lab_paired_t_test_results_within_class = d |> 
  dplyr::group_by(project, year, month, basin, class.name) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    tryCatch(
      expr = {
        t.test(.x$DB.mm3.l, .x$LV.mm3.l, paired = T) |> 
          broom::tidy() |> 
          dplyr::mutate(test_run = TRUE,
                        project = unique(.x$project), 
                        year = unique(.x$year),
                        month = unique(.x$month),
                        basin = unique(.x$basin),
                        class.name = unique(.x$class.name),
                        DB_mean = round(mean(.x$DB.mm3.l),5),
                        LV_mean = round(mean(.x$LV.mm3.l),5)) |> 
          dplyr::rename(number.rows = parameter) |> 
          dplyr::select(test_run,project,year,month,basin,class.name,DB_mean,LV_mean,dplyr::everything())},
      error = function(e) return(
        data.frame(
          test_run = FALSE,
          project = unique(.x$project), 
          year = unique(.x$year),
          month = unique(.x$month),
          basin = unique(.x$basin),
          class.name = unique(.x$class.name),
          number.rows = nrow(.x),
          DB_mean = round(mean(.x$DB.mm3.l,na.rm=T),5),
          LV_mean = round(mean(.x$LV.mm3.l,na.rm=T),5)))
    )
  },.progress=T) |> 
  dplyr::bind_rows() |> 
  tidyr::as_tibble()

# Also do this where we put all classes together inside each month.
biov_lab_paired_t_test_results = d |> 
  dplyr::group_by(project, year, month, basin) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    tryCatch(
      expr = {
        t.test(.x$DB.mm3.l, .x$LV.mm3.l, paired = T) |> 
          broom::tidy() |> 
          dplyr::mutate(test_run = TRUE,
                        project = unique(.x$project), 
                        year = unique(.x$year),
                        month = unique(.x$month),
                        basin = unique(.x$basin),
                        DB_mean = round(mean(.x$DB.mm3.l),5),
                        LV_mean = round(mean(.x$LV.mm3.l),5)) |> 
          dplyr::rename(number.rows = parameter) |> 
          dplyr::select(test_run,project,year,month,basin,DB_mean,LV_mean,dplyr::everything())},
      error = function(e) return(
        data.frame(
          test_run = FALSE,
          project = unique(.x$project), 
          year = unique(.x$year),
          month = unique(.x$month),
          basin = unique(.x$basin),
          number.rows = nrow(.x),
          DB_mean = round(mean(.x$DB.mm3.l,na.rm=T),5),
          LV_mean = round(mean(.x$LV.mm3.l,na.rm=T),5)))
    )
  },.progress=T) |> 
  dplyr::bind_rows() |> 
  tidyr::as_tibble()

abund_lab_paired_t_test_box_plots = abund_lab_paired_t_test_results_within_class |> 
  dplyr::filter(p.value < 0.05) |> 
  dplyr::group_by(project, year, month, basin, class.name) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    d |> 
      dplyr::filter(year == .x$year, month == .x$month,
                    basin == .x$basin, project == .x$project) |> 
      dplyr::select(-c(DB.mm3.l,LV.mm3.l)) |> 
      tidyr::pivot_longer(cols = c(DB.cells.ml,LV.cells.ml)) |> 
      ggplot(aes(x = class.name, y = value, fill = name)) + 
      geom_boxplot() + 
      labs(x = "Class", y = "Cell Count", fill = "Lab",
           title = paste0("Classes with Significantly Different Cell Numbers between Labs"),
           subtitle = paste0(.x$month," ",.x$year,", ",.x$basin,", ",.x$project)) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  })


biov_lab_paired_t_test_box_plots = biov_lab_paired_t_test_results_within_class |> 
  dplyr::filter(p.value < 0.05) |> 
  dplyr::group_by(project, year, month, basin, class.name) |> 
  dplyr::group_split() |> 
  purrr::map( ~ {
    d |> 
      dplyr::filter(year == .x$year, month == .x$month,
                    basin == .x$basin, project == .x$project) |> 
      dplyr::select(-c(DB.cells.ml,LV.cells.ml)) |> 
      tidyr::pivot_longer(cols = c(DB.mm3.l,LV.mm3.l)) |> 
      ggplot(aes(x = class.name, y = value, fill = name)) + 
      geom_boxplot() + 
      labs(x = "Class", y = "Biovolume", fill = "Lab",
           title = paste0("Classes with Significantly Different Biovolume between Labs"),
           subtitle = paste0(.x$month," ",.x$year,", ",.x$basin,", ",.x$project)) + 
      theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
  })
```

#### Table {.tabset}

Note 1: These T-tests compare the Abundance Biovolume recordings of Darren and Lidija by project, year, month, basin; an additional pair of tests also split th data by class name. Some of these data 'cuts' had too few rows of data to run the test (e.g. a single row): see column "number.rows" to see the total number of rows of data for both labs combined.

Note 2: percent difference of mean Biovolume is calculated as $100*(A-B)/((A+B)/2)$, where *A* is Darren's lab's values (`DB.mm3.l`) and *B* is Lidija's lab's values (`LV.mm3.l`).

##### T-Test for Classes Separately

Abundance:
```{r}
abund_lab_paired_t_test_results_within_class |> 
  dplyr::filter(!is.na(class.name)) |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

Biovolume: 
```{r}
biov_lab_paired_t_test_results_within_class |> 
  dplyr::filter(!is.na(class.name)) |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

##### T-Tests by Month (Classes Combined)

Abundance:
```{r}
abund_lab_paired_t_test_results |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

Biovolume:
```{r}
biov_lab_paired_t_test_results |> 
  dplyr::select(-test_run) |> 
  dplyr::mutate(perc_diff = paste0(round(100*(DB_mean - LV_mean)/(DB_mean + LV_mean),2),"%")) |> 
  dplyr::mutate(dplyr::across(c(estimate,statistic,p.value,conf.low,conf.high), \(x) round(x,5))) |> 
  dplyr::select(project:LV_mean,perc_diff,dplyr::everything()) |> 
  DT::datatable()
```

#### Boxplots

```{r}
abund_lab_paired_t_test_box_plots
```

```{r}
biov_lab_paired_t_test_box_plots
```

## Lab Comparison Plots {.tabset}

### Wahleach {.tabset}

#### Abundance by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Abundance by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Biovolume by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Biovolume by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Species Richness by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Species Richness by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Edibility by Month - 2023
```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Edibility by Month - 2024
```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Wahleach') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

### Alouette {.tabset}

#### Abundance by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Abundance by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.cells.ml","LV.cells.ml"),
                     var_label = "Abundance (cells*mL<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Biovolume by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Biovolume by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(vars = c("DB.mm3.l","LV.mm3.l"), 
                     var_label = "Biovolume (mm<sup>3</sup> * L<sup>-1</sup>)",
                     plot_type = 'classes_by_month',
                     t_test_results = biov_lab_paired_t_test_results)
```

#### Species Richness by Month - 2023

```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Species Richness by Month - 2024

```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Species Richness Count",
                     plot_type = 'species_richness')
```

#### Edibility by Month - 2023
```{r}
d |> 
  dplyr::filter(year == 2023, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```

#### Edibility by Month - 2024
```{r}
d |> 
  dplyr::filter(year == 2024, project == 'Alouette') |> 
  make_plot_by_month(var_label = "Edibility",
                     plot_type = 'edibility',
                     t_test_results = abund_lab_paired_t_test_results)
```
